% !TEX TS-program = lualatex
\documentclass[notoc,notitlepage]{tufte-book}
% \nonstopmode % uncomment to enable nonstopmode

\usepackage{classnotetitle}

\title{JAPAN202R S19 --- Second Year Japanese 2}
\author{Johnson Ng (ジョンソン)}
\subtitle{ノート}
\credentials{BMath (Hons), Pure Mathematics major, Actuarial Science Minor}
\institution{University of Waterloo}

\usepackage{longtable,multirow}
\usepackage{titlesec}
\usepackage{luatexja}
\usepackage{luatexja-fontspec}
\usepackage{pxrubrica}
\input{latex-classnotes-preamble.tex}

\theoremprework{\textcolor{blue}{\hrule height 2pt width \textwidth}}
\theoremheaderfont{\color{blue}\normalfont\bfseries}
\theorempostwork{\textcolor{blue}{\hrule height 2pt width \textwidth}}
\theoremindent10pt
\newtheorem{grammar}{\faBook Grammar}

\setmainfont[Ligatures=TeX]{P052}
\setmainjfont{M+ 1mn}

\begin{document}
% \input{latex-classnotes-header.tex}
\makeatletter
\fancyhead[LE]{\thepage \enspace \textsl{\leftmark}}
\makeatother

\hypersetup{pageanchor=false}
\maketitle
\hypersetup{pageanchor=true}
\begin{fullwidth}
\chaptermark{Table of Contents}
\tableofcontents
\end{fullwidth}

\makeatletter
\fancyhead[LE]{\thepage \enspace \textsl{\leftmark \enspace \rightmark}}
\makeatother

\chapter*{Preface}%
\label{chp:preface}
\addcontentsline{toc}{chapter}{Preface}
% chapter preface

This set of notes will follow almost precisely the presentation in
\citealt{banno1999}.

The notes is also meant for readers that are already somewhat familiar
with Japanese culture and speaking patterns, and so it shall be more
focused on the linguistics, rather than bringing the reader
to a level where they can understand, read and write some Japanese.
As such, the GENKI series is probably not the ideal material,
but this is also partly just to study for the exam on campus.

% chapter preface (end)

\chapter{第１８課}%
\label{chp:dai_18_ka}
% chapter dai_18_ka

\section{単語（１８課）}%
\label{sec:tango_c18}
% section tango_c18

\begin{longtable}{r l l l}
\multicolumn{4}{l}{\hlnotea{名詞}} \\
* & あと           & 後         & the rest \\
  & エアコン       &            & air conditioner \\
  & カーテン       &            & curtain \\
  & クッション     &            & cushion \\
  & シャンプー     &            & shampoo \\
* & しょうゆ       & 醤油       & soy sauce \\
* & スイッチ       &            & switch \\
  & スープ         &            & soup \\
* & スカート       &            & skirt \\
* & そと           & 外         & outside \\
  & ソファ         &            & sofa \\
* & タオル         &            & towel \\
  & にっき         & 日記       & diary \\
  & バナナ         &            & banana \\
  & ポップコーン   &            & popcorn \\
  & むし           & 虫         & insect \\
  & やちん         & 家賃       & rent \\
* & ゆうがた       & 夕方       & evening \\
  & るすばんでんわ & 留守番電話 & answering machine \\
* & れいぞうこ     & 冷蔵庫     & refrigerator \\
  & ろうそく       & 蝋燭       & candle \\
\multicolumn{4}{l}{\hlnotea{イ --- 形容詞}} \\
  & あかるい         & 明るい     & bright \\
  & きぶんがわるいう & 気分が悪い & to feel sick \\
  & くらい           & 暗い       & dark \\
  & はずかしい       & 恥ずかしい & embarrassing; \\
  &                  &            & to feel embarrassed \\
\multicolumn{4}{l}{\hlnotea{ウ --- 動詞}} \\
  & あく       & 開く       & (something) opens \\
  & あやまる   & 謝る       & to apologize \\
* & おす       & 押す       & to press; to push \\
* & おとす     & 落とす     & to drop (something) \\
  & おゆがわく & お湯が沸く & to boil water \\
  & ころぶ     & 転ぶ       & to fall down \\
  & こわす     & 壊す       & to break (something) \\
  & さく       & 咲く       & to bloom \\
  & しまる     & 閉まる     & to close (something) \\
* & たすかる   & 助かる     & to be saved; to be helped \\
* & たのむ     & 頼む       & to ask a favor \\
* & つく       &            & to turn (appliance) on \\
  & よごす     & 汚す       & to make dirty \\
\multicolumn{4}{l}{\hlnotea{ル --- 動詞}} \\
  & おちる     & 落ちる   & to drop (something) \\
* & かたづける & 片付ける & to tidy up \\
  & かんがえる & 考える   & to think (about); to consider \\
  & きえる     & 消える   & (something) goes off \\
  & こわれる   & 壊れる   & (something) breaks \\
* & よごれる   & 汚れる   & to become dirty \\
\multicolumn{4}{l}{\hlnotea{特別動詞}} \\
  & ちゅうもんする & 注文する & to place an order \\
\multicolumn{4}{l}{\hlnotea{他の}} \\
* & いますぐ               & 今すぐ             & right away \\
* & おかげで               &                    & thanks to ... \\
* & おさきにしつれいします & お先に失礼します   & I shall be leaving first \\
* & おつかれさま（でした） & お疲れ様（でした） & That must be tiring for you \\
* & 〜（ん）だろう         &                    & short for 〜（ん）でしょう \\
* & どうしよう             &                    & What should I/we do? \\
* & ほんとうに             & 本当に             & really \\
* & まず                   &                    & first of all \\
  & 〜までに               &                    & by (time/date)
\end{longtable}

% section tango_c18 (end)

\section{自動詞と他動詞}%
\label{sec:jidoushi_to_tadoushi}
% section jidoushi_to_tadoushi

This is a direct equivalence of \hlnotea{transitivity} in English, where
\begin{itemize}
  \item 自動詞 --- intransitive verbs
  \item 他動詞 --- transitive verbs
\end{itemize}

\begin{grammar}[自動詞]\index{自動詞}\label{grammar:jidoushi}
  \hlnoteb{自動詞} (intransitive verbs) are verbs that describe the changes or
  the state that things or people undergo.
\end{grammar}

\begin{grammar}[他動詞]\index{他動詞}\label{grammar:tadoushi}
  \hlnoteb{他動詞} (transitive verbs) are verbs that describe
  situations in which human beings act on things.
\end{grammar}

Most verbs do not have a counterpart of the opposite transitivity.
\cref{table:transitive_intransitive_verb_pair} shows some examples.

\begin{table}[ht]
  \centering
  \caption{自動詞と他動詞ペア}
  \label{table:transitive_intransitive_verb_pair}
  \begin{tabular}{l | l}
    他動詞 & 自動詞 \\
    \hline
    開ける & 開く \\
    閉める & 閉まる \\
    入れる & 入る \\
    出す   & 出る \\
    つける & つく \\
    消す   & 消える \\
    壊す   & 壊れる \\
    汚す   & 汚れる \\
    落とす & 落ちる \\
    沸かす & 沸く
  \end{tabular}
\end{table}

\begin{note}
  \begin{enumerate}
    \item 他動詞の場合はよく物や人に使うから、よく「を」と使います。 \\
      例：たけしさんが電気\underline{を}つけました。
    \item 自動詞は物を形容するみたいなことなので、よく「が」と
      使っています。 \\
      例：電気\underline{が}付きました。
  \end{enumerate}
\end{note}

There is also a subtle note that should be noted about the tenses
of transitive and intransitive verbs.

\begin{figure*}[ht]
  \centering
  \begin{tikzpicture}
    \draw[->] (-2, 0) -- (2, 0) node[right] {time};
    \draw[thick,color=green]
      (-2, -1) -- (-1, -1) node[color=foreground,midway,below] {will happen} --
      (-0.5, 1) -- (0.5, 1) node[color=foreground,midway,above] {happening} --
      (1, -1) -- (2, -1) node[color=foreground,midway,below] {happened};
    \node[below=15pt,color=green] at (-1.5, -1) {壊します};
    \node[above=15pt,color=green] at (0, 1) {壊しています};
    \node[below=15pt,color=green] at (1.5, -1) {壊しました};

    \draw[->] (4, 0) -- (8, 0) node[right] {time};
    \draw[thick,color=green]
      (4, -1) -- (5.5, -1) node[color=foreground,midway,below] {old state} --
      (6.5, 1) -- (8, 1) node[color=foreground,midway,above] {new state};
    \node[circle,fill=green,inner sep=1.5pt,
      pin={[pin edge={latex'-}, pin distance=10pt]315:{state change}}]
      at (6, 0) {};
    \node[below=15pt,color=green] at (4.75, -1) {壊れます};
    \node[below=5pt,color=green] at (7.5, -0.5) {壊れました};
    \node[above=15pt,color=green] at (7.5, 1) {壊れています};
  \end{tikzpicture}
  \caption{Tenses Subtleties with Transitive and Intransitive Verbs}
  \label{fig:tenses_subtleties_with_transitive_and_intransitive_verbs}
\end{figure*}

% section jidoushi_to_tadoushi (end)

\section{〜てしまう}%
\label{sec:_teshimau}
% section _teshimau

\begin{grammar}[〜てしまう]\index{〜てしまう}\label{grammar:teshimau}
  \hlnoteb{〜てしまう} has 2 uses:
  \begin{enumerate}
    \item to say that one ``carries out with determination''
      a plan described by the verb;
    \item to say that there is a ``lack of premeditation
      or control over how things have turned out''.
      要するに、「〜てしまう」を使うときは、
      自分がコントロールできないことや
      起こりたくないことを起こったなど
      の無力の感情を表します。
  \end{enumerate}
\end{grammar}

\begin{eg}
  To show the use of 〜しまう, we have the following examples:
  \begin{enumerate}
    \item 本を読んでしまいました。 I ended up reading the book.
    \item 電車の中にかばんを忘れてしまいました。 I left my bag in the train.
    \item 宿題を忘れたので、先生は怒ってしまいました。 I was reprimanded
      for forgetting about my homework.
  \end{enumerate}
\end{eg}

\begin{note}
  In speech, we often hear 〜てしまう and 〜でしまう contracted as
  〜ちゃう and 〜じゃう, respectively.
\end{note}

\begin{warning}
  Note that しまう goes with the verbal テ form,
  which is affirmative, and so we cannot use 〜てしまう
  to express negated ideas or actions, such as, "unfortunately,
  I did not do $x$."
\end{warning}

% section _teshimau (end)

\section{〜と}%
\label{sec:_to}
% section _to

\begin{grammar}[〜と]\index{〜と}\label{grammar:to}
  Usually used as
  \begin{center}
    A とB,
  \end{center}
  where A is a situation, and B is the consequent effect.
  〜と is one of the conditional conjugates.
\end{grammar}

\begin{note}
  The use of と here is not to be confused with its other use as an equivalence
  of ``and'' in English.
\end{note}

\begin{eg}
  \begin{enumerate}
    \item 私はその人と話すと元気になる。
    \item メアリーさんが国に帰ると寂しくなります。
    \item 秋になると木が赤くなります。
  \end{enumerate}
\end{eg}

\begin{warning}
  The event described by the second clause must be a rather clear consequence
  of the first situation.
  For instance, we do not say
  \begin{center}
    私はその人と話すと\ruby[j]{喫茶店}{きっ|さ|てん}に行きます。
  \end{center}
\end{warning}

% section _to (end)

\section{〜ながら}%
\label{sec:_nagara}
% section _nagara

\begin{grammar}[〜ながら]\index{〜ながら}\label{grammar:nagara}
  We write
  \begin{center}
    A \hlnoteb{ながら} B をします
  \end{center}
  to express that we do B while A is either being done or ongoing.
\end{grammar}

\begin{note}
  It is important to note that the two verbs that flank ながら
  must be two actions performed by the same person.

  For the case where person 2 does something
  while person 1 is doing something else, we shall use
  \nameref{sec:aidani}.
\end{note}

\begin{eg}[Basic examples for 〜ながら]
  \begin{enumerate}
    \item 私はいつもラジオ番組を聞きながら
      勉強します。
    \item たけしさんは歌を歌いながら
      掃除します。
    \item アルバイトをしながら
      学校に行くのは大変です。
  \end{enumerate}
\end{eg}

% section _nagara (end)

\section{〜ばよかったです}%
\label{sec:_bayokatsutadesu}
% section _bayokatsutadesu

\begin{grammar}[〜ばよかったです]\index{〜ばよかったです}\label{grammar:_bayokatsutadesu}
  We say
  \begin{center}
    X すれ\hlnoteb{ばよかったです}
  \end{center}
  to indicate that \hlnotef{we hope we had did X}.
  It is used to describe an alternative course of actions that we
  regret not taking.
\end{grammar}

\begin{note}
  For the rules of conjugating for 〜ば, we shall refer the reader to
  \nameref{sec:ba}, which gives a broader treatment to the form.
\end{note}

\begin{eg}
  \begin{enumerate}
    \item あの時、「愛している」と言えばよかったです。
    \item 彼女と別れなければよかったです。
  \end{enumerate}
\end{eg}

% section _bayokatsutadesu (end)

% chapter dai_18_ka (end)

\chapter{第１９課}%
\label{chp:dai_19_ka}
% chapter dai_19_ka

\section{単語（１９課）}%
\label{sec:tango_c19}
% section tango_c19

\begin{longtable}{r l l l}
\multicolumn{4}{l}{\hlnotea{名詞}} \\
* & おくさま       & 奥様           & (your/his) wife (polite) \\
  & おこさん       & お子さん       & (your/their) child (polite) \\
  & おれい         & お礼           & expression of gratitude \\
  & けいご         & 敬語           & honorific language \\
* & こちら         &                & this way (polite) \\
* & しゅっちょう   & 出張           & business trip \\
  & しゅるい       & 種類           & a kind; a sort \\
  & せいかく       & 性格           & personality; temperament \\
  & ちゅうがくせい & 中学生         & junior high school student \\
  & どちら         &                & where (polite) \\
  & なまけもの     & 怠け者         & lazy person \\
  & なやみ         & 悩み           & worry \\
  & はずかしがりや & 恥ずかしがりや & shy person \\
* & はなし         & 話             & chat; talk \\
* & ぶちょう       & 部長           & department manager; club president \\
  & ぶんか         & 文化           & culture \\
  & まちがい       & 間違い         & mistake \\
\multicolumn{4}{l}{\hlnotea{イ --- 形容詞}} \\
  & なかがいい & 仲がいい & to be on good terms; to get along well \\
\multicolumn{4}{l}{\hlnotea{ナ --- 形容詞}} \\
  & まじめな & 真面目な & serious; sober; diligent \\
\multicolumn{4}{l}{\hlnotea{ウ --- 動詞}} \\
  & いらっしゃる     &              & honorific form for いく, くる and いる \\
* & おくる           & 送る         & to walk/drive (someone) \\
  & おこる           & 怒る         & to gete angry \\
  & おっしゃる       &              & honorific form for 言う \\
* & おやすみになる   & お休みになる & honorific form for 寝る \\
* & きまる * 決まる  &              & to make a decision \\
  & くださる         & 下さる       & honorific form for くれる \\
  & ごらんいなる     & ご覧になる   & honorific form for 見る \\
  & しりあう         & 知り合う     & to get acquainted with \\
  & 〜ていらっしゃる &              & honorific form for 〜ている \\
  & なさる           &              & honorific form for する \\
  & ひっこす         & 引っ越す     & to move (to another place to live) \\
* & めしあがる       & 召し上がる   & honorific form for 食べる and 飲む \\
* & よぶ             & 呼ぶ         & to call (someone's name); to invite (person) \\
* & よる             & 寄る         & to stop by \\
\multicolumn{4}{l}{\hlnotea{ル --- 動詞}} \\
* & おくれる & 遅れる & to become late (for something) \\
  & かける   &        & to sit down (\textit{seat} に) \\
  & はれる   & 晴れる & to become sunny \\
  & もてる   &        & to be popular (in terms of romantic interest) \\
\multicolumn{4}{l}{\hlnotea{特別動詞}} \\
* & えんりょする & 遠慮する & to hold back for the time being; \\
  & & & to refrain from \\
* & ごちそうする & & to treat/invite (someone) to a meal \\
  & しょうたいする & 招待する & to invite someone (to an event/place) \\
  & ちゅういする & 注意する & to watch out; to give warning \\
* & はなしをする & 話をする & to have a talk \\
\multicolumn{4}{l}{\hlnotea{他の}} \\
  & おととい         &                & the day before yesterday \\
  & それで           &                & then; therefore \\
  & なぜ             &                & why (＝どうして) \\
* & ほんとうは       & 本当は         & in fact; originally so \\
  & まいあさ         & 毎朝           & every morning \\
* & まだ             &                & still \\
* & 〜めいさま       & 〜名様         & party of ... people \\
  & ようこそ         &                & Welcome \\
* & よろしくおつたえ & よろしくお伝え & Please give my regards (to... ) \\
  & ください         & ください       &
\end{longtable}

% section tango_c19 (end)

\section{尊敬語 Honorific Verbs}%
\label{sec:sonkeigo_honorific_verbs}
% section sonkeigo_honorific_verbs

We use honorific verbs to show respect to the person
of which the action pertains to.

The following are some of the verbs converted to their
honorific forms, along with their \hlnotec{irregular conjugations}.

\begin{table}[ht]
  \centering
  \caption{Table of Basic Verbs in their Honorific Form and Their Irregular Conjugation}
  \label{table:table_of_basic_verbs_in_their_honorific_form_and_their_irregular_conjugation}
  \begin{tabular}{l | l l}
    Basic Verb & Honorific Verbs             & Irregular Conjugation \\
    \hline
    いる       &                             & \\
    行く       & いらっしゃる                & いらしゃ\underline{い}ます \\
    来る       &                             & \\
    \hline
    見る       & ご覧になる                  & \\
    \hline
    言う       & おっしゃる                  & おっしゃ\underline{い}ます \\
    \hline
    する       & なさる                      & なさ\underline{い}ます \\
    \hline
    食べる     & \multirow{2}{*}{召し上がる} & \\
    飲む       &                             & \\
    \hline
    くれる     & くださる                    & くださ\underline{い}ます \\
    \hline
    寝る       & お休みになる                & \\
    \hline
    〜ている   & 〜ていらっしゃる            & 〜ていらっしゃ\underline{い}ます
  \end{tabular}
\end{table}

\begin{warning}
  Honorific verbs are directed towards describing the action
  of the person that we speak of.
  We never use honorific verbs to describe our own actions.
\end{warning}

\begin{eg}
  \begin{enumerate}
    \item 先生は今日学校にいらっしゃいません。
    \item 何が召し上がりますか。
    \item 田中さんのお母さんがこの本をくださいました。
    \item 先生は自分で料理なさるそうです。
  \end{enumerate}
\end{eg}

\begin{note}
  For verbs that lack an honorific counterpart,
  we add the respect factor as follows:
  \begin{enumerate}
    \item Using ていらっしゃいます instead of ています,
      if the sentence has the helping verb ている. \\
      例：先生が電話で話していらっしゃいます。\\
      例：先生は疲れていらっしゃるみたいです。
    \item Flank the verb stem with お and になる,
      in most cases. \\
      例：先生はもうお帰りになりました。 \\
      例：この雑誌をお読みになったことがありますか。
  \end{enumerate}
\end{note}

\begin{note}
  As shown in the above examples,
  we can turn most combinations of a verb and a post-predicate expression
  into the honorific style simply by turning the verb into the honorific form.
  However, post-predicate expressions such as
  ことがあります and ください remain unchanged.
  This also applies to expressions like てもいい and てはいけない,
  to the potential verbs.

  It is noteworthy that it is note in good taste to talk about what
  the respected person can or cannot do, may or must not do.

  ている is exceptional in being a post-predicate that regularly undergoes the
  honorific style change.
  Special honorific verbs generally \hlnoted{take priority} over
  ていらっしゃる, such as in
  \begin{center}
    先生はテレビを\hlnoted{ご覧になっています}。
  \end{center}
  However, forms like 見ていらっしゃいます are also acceptable.
\end{note}

% section sonkeigo_honorific_verbs (end)

\section{Giving Respectful Advice}%
\label{sec:giving_respectful_advice}
% section giving_respectful_advice

We usually hear the form
\begin{center}
  お + verb stem + ください
\end{center}
in public address announcements and in the speech of store attendants.

\begin{eg}
  \begin{enumerate}
    \item 切符をお取りください。
    \item 説明をお読みください。
  \end{enumerate}
\end{eg}

While the sentences end with ください,
it is best to consider them as courteously phrased commands,
rather than a request.
In particular, you are `advised' to perform said action for your own sake.
Thus, it is inappropriate to say
\begin{center}
  $\times$ \ruby[j]{塩}{しお}をお取りください。
\end{center}
to mean that you wish for someone to pass the salt to you.

\begin{eg}
  Some further examples of giving respectful advice are the following:
  \begin{enumerate}
    \item ご注意ください。
    \item ご覧ください。
    \item お召し上がりください。
    \item お休みください。
  \end{enumerate}
  Notice the use of ご instead of お with certain compound verbs.
\end{eg}

% section giving_respectful_advice (end)

\section{〜てくれてありがとう}%
\label{sec:_tekuretearigatou}
% section _tekuretearigatou

\begin{grammar}[〜てくれてありがとう]\index{〜てくれてありがとう}\label{grammar:_tekuretearigatou}
  When we want to \hlnotef{express gratitude to someone for a specific action},
  we can say
  \begin{center}
    verb テ form + くれてありがとう.
  \end{center}
\end{grammar}

\begin{eg}
  \begin{enumerate}
    \item 手伝ってくれてありがとう。
    \item 推薦状を書いてくださってありがとうございました。
    \item いい友達でいてくれてありがとう。
  \end{enumerate}
\end{eg}

% section _tekuretearigatou (end)

\section{〜てよかったです}%
\label{sec:_teyokatsutadesu}
% section _teyokatsutadesu

\begin{grammar}[〜てよかった]\index{〜てよかった}\label{grammar:_teyokatsuta}
  We say
  \begin{center}
    verb テ form てよかったです
  \end{center}
  to express that we are \hlnotef{glad that a certain action (or inaction) is done}.
\end{grammar}

\begin{eg}
  \begin{enumerate}
    \item 日本語を勉強してよかったです。
    \item メアリーさんが元気になってよかったです。
    \item きのうのパーティーに行かなくてよかったです。
  \end{enumerate}
\end{eg}

% section _teyokatsutadesu (end)

\section{〜はずです}%
\label{sec:_hazudesu}
% section _hazudesu

\begin{grammar}[〜はずです]\index{〜はずです}\label{grammar:_hazudesu}
  We say
  \begin{center}
    verb stem + はずです
  \end{center}
  to show that an action was \hlnotef{supposed to be the case}.

  We may also use 〜はずです with adjectives and nouns,
  almost without explicit modification but in their modal forms.
\end{grammar}

\begin{eg}
  \begin{enumerate}
    \item 今日は日曜日だから、銀行は閉まっているはずです。
    \item きのうメアリーさんはどこにも行かなかったはずです。
    \item この試験は難しいはずです。
    \item そんなはずじゃなかったです。
    \item 日本人のはずです。
  \end{enumerate}
\end{eg}

% section _hazudesu (end)

% chapter dai_19_ka (end)

\chapter{第２０課}%
\label{chp:dai_20_ka}
% chapter dai_20_ka

\section{単語（２０課）}%
\label{sec:tango_c20}
% section tango_c20

\begin{longtable}{r l l l}
\multicolumn{4}{l}{\hlnotea{名詞}} \\
  & あちら       &          & that way (polite) \\
  & アニメ       &          & animation \\
  & うちゅうじん & 宇宙人   & space alien \\
* & おと         & 音       & sound \\
  & おにぎり     &          & rice ball \\
* & かかりのもの & 係の者   & (our) person in charge \\
* & かど         & 角       & corner \\
  & くうこう     & 空港     & airport \\
  & じ           & 字       & letter; character \\
  & してん       & 支店     & branch office \\
  & しゅみ       & 趣味     & hobby; pastime \\
  & しょうせつ   & 小説     & novel \\
  & しんご       & 信号     & traffice light \\
  & スニーカー   &          & sneakers \\
* & せんす       & 扇子     & paper fan \\
  & つき         & 月       & moon \\
* & でんしじしょ & 電子辞書 & electronic dictionary \\
  & ドイツ       &          & Germany \\
  & ハイヒール   &          & high heels \\
* & 〜や         & 〜屋     & .. shop \\
\multicolumn{4}{l}{\hlnotea{イ --- 形容詞}} \\
* & おもい & 重い & heavy; serious (illness, topic) \\
  & かるい & 軽い & light \\
\multicolumn{4}{l}{\hlnotea{ウ --- 動詞}} \\
* & いたす     & 致す   & extra-modest form for する \\
* & いただく   & 頂く   & extra-modest form for 食べる and 飲む; \\
  &            &        & humble form for もらう \\
  & うかがう   & 伺う   & to humbly visit; to humbly ask \\
  & おる       &        & extra-modest form for いる \\
* & ござる     &        & extra-modest form for ある \\
  & 〜ておる   &        & extra-modest form for 〜ている \\
  & 〜てござる &        & extra-modest form for です \\
* & まいる     & 参る   & extra-modest form for いく and くる \\
* & まがる     & 曲がる & to run at a corner \\
* & もうす     & 申す   & extra-modest form for 言う \\
  & もどる     & 戻る   & to come back \\
\multicolumn{4}{l}{\hlnotea{ル --- 動詞}} \\
* & きこえる   & 聞こえる   & to be audible \\
  & さしあげる & 差し上げる & humble form for あげる \\
  & つたえる   & 伝える     & to convey (a message) \\
* & またせる   & 待たせる   & to keep (someone) waiting \\
\multicolumn{4}{l}{\hlnotea{特別動詞}} \\
* & こうかんする & 交換する & to exchange \\
  & せいかつする & 生活する & to lead a life \\
* & へんぴんする & 返品する & to return (merchandise) \\
\multicolumn{4}{l}{\hlnotea{他の}} \\
* & おや？           &                  & Oh? \\
  & 〜かい           & 〜階             & ...th floor \\
* & かしこまりました &                  & Certainly. \\
  & さあ             &                  & I am not very sure, ... \\
* & しつれいしました & 失礼いたしました & I am very sorry. \\
* & しょうしょう     & 少々             & a few seconds \\
  & それでは         &                  & if that is the case, ... \\
* & できれば         &                  & if possible \\
* & まことに         & 誠に             & really; sincerely (very polite) \\
  & また             &                  & again \\
  & 〜みたいな       &                  & such as... \\
* & もうしわけ       & 申し訳           & You have my apologies. \\
  & ありません       & ありません       & \\
* & よろしかったら   &                  & if it is okay (polite)
\end{longtable}

% section tango_c20 (end)

\section{丁重語 Extra-modest Expressions}%
\label{sec:teichougo_extra_modest_expressions}
% section teichougo_extra_modest_expressions

To speak modestly of our own actions, we shall use what is called
\ruby[j]{丁重語}{てい|ちょう|ご}.
\cref{table:basic_extra_modest_expressions} shows some of the basic
extra-modest expressions.

\begin{table}[ht]
  \centering
  \caption{Basic Extra-modest Expressions}
  \label{table:basic_extra_modest_expressions}
  \begin{tabular}{l l l}
             & \multicolumn{2}{l}{extra-modest expression} \\
    いる     & おります                      & (おる) \\
    行く     & \multirow{2}{*}{参ります}     & \multirow{2}{*}{(参る)} \\
    来る     &                               & \\
    言う     & 申します                      & (申す) \\
    する     & いたします                    & (いたす) \\
    食べる   & \multirow{2}{*}{いただきます} & \multirow{2}{*}{(いただく)} \\
    飲む     &                               & \\
    ある     & ございます                    & (ござる) \\
    〜ている & 〜ております                  & (〜ておる) \\
    〜です   & 〜でございます                & (〜でござる)
  \end{tabular}
\end{table}

\begin{eg}
  \begin{enumerate}
    \item 私は来年も日本におります。 cf. います
    \item 私は今年の六月に大学を卒業いたしました。 cf. 卒業します
    \item 電車が参ります。
    \item 父が家におります。
  \end{enumerate}
\end{eg}

% section teichougo_extra_modest_expressions (end)

\section{謙虚語 Humble Expressions}%
\label{sec:kenkyogo_humble_expressions}
% section kenkyogo_humble_expressions

When doing something out of respect for somebody,
we can describe our actions using a verb in the humble pattern
\begin{center}
  お + stem + する.
\end{center}
Note that certain する compound verbs do not follow this pattern,
but instead takes on the form
\begin{center}
  ご + base compound verb + する.
\end{center}

\begin{eg}
  \begin{enumerate}
    \item 私は昨日先生にお会いしました。
    \item 私は先生の本をお貸しするつもりです。
    \item 部長にお電話します。
    \item お客様をご案内します。
  \end{enumerate}
\end{eg}

\begin{warning}
  It has been debated that we should avoid using 差し上げる
  to say that ``we do something for somebody''.
  The argument is that the idea of doing something for somebody
  as a service is but an insolent belief, and having to speak
  of it in a humble way is an unconvincing facade.
  Instead, we make use of the お + stem + する form in this cases.
\end{warning}

% section kenkyogo_humble_expressions (end)

\section{〜ないで}%
\label{sec:_naide}
% section _naide

\begin{grammar}[〜ないで]\index{〜ないで}\label{grammar:_naide}
  When doing something without doing something else,
  we say
  \begin{center}
    X し\hlnoteb{ないで} Y します
  \end{center}
  to mean that we do Y without doing X.
\end{grammar}

\begin{note}
  We use 〜ない for both present and past actions.
  One may think of the speaker say that they did not do X
  while thinking of themselves as being at that time when
  the action was expected to happen.
\end{note}

\begin{eg}
  \begin{enumerate}
    \item 昨日の夜は、寝ないで、勉強しました。
    \item 辞書を使わないで、新聞を読みます。
  \end{enumerate}
\end{eg}

% section _naide (end)

\section{Questions within Larger Sentences}%
\label{sec:questions_within_larger_sentences}
% section questions_within_larger_sentences

One can include a question as part of a longer sentence
and express ideas such as ``I don't know when the test is''
and ``I don't remember whether Mary came to the party.''

\begin{eg}
  We shall highlight the quoted question clauses in the following
  examples:
  \begin{enumerate}
    \item 山下先生は\hlnoted{昨日何を食べたか}覚えていません。
    \item \hlnoted{メアリーさんは来たかどうか}知っていますか。 
  \end{enumerate}
\end{eg}

% section questions_within_larger_sentences (end)

\section{name という item}%
\label{sec:name_toiu_item}
% section name_toiu_item

\begin{grammar}[name という item]\index{name という item}\label{grammar:name_toiu_item}
  When we talk about something that we do not think our
  listener may not be familiar with, we can use the pattern
  \begin{center}
    name という category,
  \end{center}
  where the category tries to describe the item in the name
  usually in a generic sense.
\end{grammar}

\begin{eg}
  \begin{enumerate}
    \item ポチという犬
    \item 「花」という歌
  \end{enumerate}
\end{eg}

% section name_toiu_item (end)

\section{〜やすい / 〜にくい}%
\label{sec:_yasui_nikui}
% section _yasui_nikui

When describing that something is ``easy to do'' or
``difficult to do'', we use the conjugatation
〜やすい and 〜にくい, respectively.

\begin{eg}
  \begin{enumerate}
    \item この電子辞書は使いやすいです。 cf 使う
    \item 骨が多いので、魚は食べにくいです。 cf 食べる
  \end{enumerate}
\end{eg}

% section _yasui_nikui (end)

% chapter dai_20_ka (end)

\chapter{第２１課}%
\label{chp:dai_21_ka}
% chapter dai_21_ka

\section{単語（２１課）}%
\label{sec:tango_c21}
% section tango_c21

\begin{longtable}{r l l l}
\multicolumn{4}{l}{\hlnotea{名詞}} \\
  & あかちゃん & 赤ちゃん & baby \\
  & か         & 蚊       & mosquito \\
  & かいぎ     & 会議     & business meeting; conference \\
  & ガソリン   &          & gasoline \\
  & かんきょう & 環境     & environment \\
* & けいさつ   & 警察     & police \\
  & こうじょう & 工場     & factory \\
* & こと       & 事       & things; matters \\
* & しゅうでん & 終電     & last train \\
  & じゅんび   & 準備     & preparation \\
  & スピーチ   &          & speech \\
  & せいふ     & 政府     & government \\
  & ちかん     &          & sexual offender; prevent \\
  & どうりょう & 同僚     & colleague \\
* & どろぼう   & 泥棒     & thief; burglar \\
* & バイト     &          & short for アルバイト \\
* & はんにん   & 犯人     & criminal \\
  & ポスター   &          & poster \\
  & むかし     & 昔       & old days; past \\
  & もんく     & 文句     & complaint \\
* & るす       & 留守     & absence; not at home \\
\multicolumn{4}{l}{\hlnotea{イ --- 形容詞}} \\
  & とおい & 遠い & far away \\
  & ひどい & 酷い & awful; cruel \\
\multicolumn{4}{l}{\hlnotea{ナ --- 形容詞}} \\
  & あんぜん     & 安全 & safe \\
  & たいせつ     & 大切 & precious; valuable \\
* & めちゃくちゃ &      & messy; disorganized \\
\multicolumn{4}{l}{\hlnotea{ウ --- 動詞}} \\
  & おく         & 置く       & to put; to lay; to place \\
* & きがつく     & 気が付く   & to notice \\
  & ける         &            & to kick \\
  & さす         & 刺す       & to sting; to poke \\
  & さわる       & 触る       & to touch \\
* & つかまる     & 捕まる     & to be arrested; to be caught \\
  & つつむ       & 包む       & to wrap; to cover \\
  & なぐる       & 殴る       & to strike; to hit; to punch \\
  & ぬすむ       & 盗む       & to steal; to rob \\
  & はる         & 貼る       & to post; to stick \\
  & ふむ         & 踏む       & to step on \\
  & ふる         &            & to turn down; to reject; to jilt \\
  & もんくをいう & 文句を言う & to complain \\
  & やく         & 焼く       & to bake; to burn; to grill \\
  & やる         &            & to give (to pets, plants, younger siblings, etc.) \\
\multicolumn{4}{l}{\hlnotea{ル --- 動詞}} \\
  & いじめる   &          & to bully \\
  & きがえる   & 着替える & to change clothes \\
* & ためる     &          & to save money \\
  & つづける   & 続ける   & to continue \\
  & ほめる     & 褒める   & to praise; to say nice things \\
  & まちがえる & 間違える & to make a mistake \\
  & みつける   & 見つける & to find \\
\multicolumn{4}{l}{\hlnotea{特別動詞}} \\
  & ばかにする   &            & to insult; to make a fool of... \\
* & びっくりする &            & to be surprised \\
  & ひるねをする & 昼寝をする & to take a nap \\
* & れんらくする & 連絡する   & to contact \\
\multicolumn{4}{l}{\hlnotea{他の}} \\
* & 〜あいだに & 〜間に & while ... \\
* & ころ       & 頃     & time of ...; when ... \\
  & すこし     & 少し   & a little \\
* & とにかく   &        & anyway; anyhow
\end{longtable}

% section tango_c21 (end)

\section{受動文 Passive Sentences}%
\label{sec:jyudoubun_passive_sentences}
% section jyudoubun_passive_sentences

We can change verbs into their passive forms, and use them as we do in English.

\begin{table}[ht]
  \centering
  \caption{Rules of converting from base form to passive form}
  \label{table:rules_of_converting_from_base_form_to_passive_form}
  \begin{tabular}{l c l c l c l}
  \multicolumn{7}{l}{ル --- 動詞: 「る」は「られる」になる} \\
  食べる & $\to$ & 食べられる &  &  &  &  \\
  $  $ \\
  \multicolumn{7}{l}{う --- 動詞: 最後の「う」の音は「あれる」になる} \\
  行く & $\to$ & 行かれる & & 話す & $\to$ & 話される \\
  待つ & $\to$ & 待たれる & & 死ぬ & $\to$ & 死なれる \\
  読む & $\to$ & 読まれる & & 取る & $\to$ & 取られる \\
  泳ぐ & $\to$ & 泳がれる & & 遊ぶ & $\to$ & 遊ばれる \\
  買う & $\to$ & 買われる \\
  $  $ \\
  \multicolumn{7}{l}{特別動詞} \\
  来る & $\to$ & 来られる & & する & $\to$ & される
  \end{tabular}
\end{table}

\begin{note}
  One may notice that the passive forms for ル動詞 and 特別動詞
  are the same as their potential forms, while
  ウ動詞 are different from their potential counterparts.
\end{note}

% section jyudoubun_passive_sentences (end)

\section{〜てある}%
\label{sec:tearu}
% section tearu

\begin{grammar}[〜てある]\index{〜てある}\label{grammar:_tearu}
  We use the テ form of a verb, along with the helping verb ある
  to characterize a situation that \hlnotee{has been brought about
  on purpose} by somebody who may be unnamed or unknown.

  We may also use 〜てあります if somebody, possibly ourselves,
  performed an action on purpose earlier, which can be described
  in terms of 〜ておきました, and if the result of the action
  is still observable at the moment.
\end{grammar}

\begin{note}
  Since 〜てある describes the state of an object, we typically use
  the particle が for the subject.
\end{note}

\begin{eg}
  \begin{enumerate}
    \item 寒いので、ヒーターがつけてあります。
    \item テーブルの上に本がおいてあります。
    \item レストランの予約がしてあります。 as a result of
      予約をしておきました。
    \item パンが買ってあります。 as a result of
      パンを買っておきました。
  \end{enumerate}
\end{eg}

% section tearu (end)

\section{〜間に}%
\label{sec:aidani}
% section aidani

\begin{grammar}[〜間に]\index{〜間に}\label{grammar:_aidani}
  We use the pattern ``A 間に B'' when the event B takes place 
  in the \hlnotec{middle} of the occurrence of A, which is an
  event that has a certain duration.
\end{grammar}

\begin{note}[Difference between 間 and 間に]
  ``A 間に B'' is reserved for activities (B) that occur within the time period
  of which activity A is ongoing.
  \begin{marginfigure}
    \centering
    \begin{tikzpicture}
      \draw (-2, 0) -- (2, 0);
      \draw[{Bracket[length=5pt,width=15pt]}-{Bracket[length=5pt,width=15pt]},red]
        (-1.5, 0) -- (1.5, 0);
      \draw[<->,red] (-1.5, -0.5) -- (1.5, -0.5)
        node[midway,below] {Duration of A};
      \draw[{Bracket[length=5pt,width=15pt]}-{Bracket[length=5pt,width=15pt]},blue]
        (-0.5, 0) -- (0.5, 0);
      \draw[<->,blue] (-0.5, 0.5) -- (0.5, 0.5)
        node[midway,above] {Duration of B};
    \end{tikzpicture}
    \caption{Rule for 〜間に}\label{fig:rule_for_aidani}
  \end{marginfigure}

  ``A 間 B'' is for cases where B happens throughout the time at which A occurs.
  \begin{marginfigure}
    \centering
    \begin{tikzpicture}
      \draw (-2, 0) -- (2, 0);
      \draw[{Bracket[length=5pt,width=15pt]}-{Bracket[length=5pt,width=15pt]},red]
        (-1.5, 0) -- (1.5, 0);
      \draw[<->,red] (-1.5, -0.5) -- (1.5, -0.5)
        node[midway,below] {Duration of A and B};
    \end{tikzpicture}
    \caption{Rule for 〜間}\label{fig:rule_for_aida}
  \end{marginfigure}
\end{note}

\begin{eg}
  \begin{enumerate}
    \item お風呂に入っている間に電話があります。
    \item 昨日の夜、寝ている間に地震がありました。
    \item 留守の間に友達が来ました。
  \end{enumerate}
\end{eg}

% section aidani (end)

\section{形容詞＋する}%
\label{sec:keiyoushi_suru}
% section keiyoushi_suru

\begin{grammar}[形容詞＋する]\index{形容詞＋する}\label{grammar:keiyoushi_suru}
  We use （形容詞）＋する to mean we make the described adjective happen.
\end{grammar}

\begin{eg}
  \begin{enumerate}
    \item 寒くにする : to make something cold/colder
    \item 簡単にする : to make something easy/easier
    \item この間の試験は難しすぎたので、次の試験は優しくしてください。
    \item 部屋をきれいにしました。
  \end{enumerate}
\end{eg}

% section keiyoushi_suru (end)

\section{〜てほしい}%
\label{sec:tehoshii}
% section tehoshii

\begin{grammar}[〜てほしい]\index{〜てほしい}\label{grammar:_tehoshii}
  When we want somebody to do something, recall that in English,
  we can say that ``I wish you/he/she would...''.
  There is an equivalent form in Japanese, in the form of 〜てほしい.
\end{grammar}

\begin{note}[Particle note]
  The person of which the wish is directed to is marked with the particle
  に.
\end{note}

\begin{eg}
  \begin{enumerate}
    \item 私は病気の友達に元気になってほしいです。
    \item ルームメートに宿題を手伝ってほしかったです。
    \item お父さんに昔の話をしてほしくないです。
    \item ホストファミリーに英語で話さないでほしいです。
  \end{enumerate}
\end{eg}

% section tehoshii (end)

% chapter dai_21_ka (end)

\chapter{第２２課}%
\label{chp:dai_22_ka}
% chapter dai_22_ka

\section{単語（２２課）}%
\label{sec:tango_c22}
% section tango_c22

\begin{longtable}{r l l l}
\multicolumn{4}{l}{\hlnotea{名詞}} \\
  & あいて       & 相手       & partner / opponent \\
  & うけつけ     & 受付       & reception desk \\
* & えいかいわ   & 英会話     & English conversation \\
  & おじょうさん & お嬢さん   & (someone's) daughter (polite) \\
  & かじ         & 家事       & household matters, housework \\
  & かぜ         & 風         & wind \\
  & かみさま     & 神様       & God \\
  & き           & 木         & wood / tree \\
  & こうはい     & 後輩       & junior member of a group \\
  & さる         & 猿         & monkey \\
* & じゆう       & 自由       & freedom \\
* & じゅく       & 塾         & cram school \\
  & しょるい     & 書類       & document \\
  & せんぱい     & 先輩       & senior member of a group \\
  & ひとりぐらし & 一人暮らし & living alone \\
  & ぶか         & 部下       & subordinate \\
  & ふくしゅう   & 復習       & revision \\
  & プロジェクト &            & project \\
  & ボール       &            & ball \\
  & むだづかい   & 無駄遣い   & waste (money) \\
  & めんきょ     & 免許       & license \\
  & ヨーロッパ   &            & Europe \\
  & よしゅう     & 予習       & preview of lessons \\
\multicolumn{4}{l}{\hlnotea{イ --- 形容詞}} \\
* & うるさい & & noisy; annoying \\
\multicolumn{4}{l}{\hlnotea{ナ --- 形容詞}} \\
* & しんぱい & 心配 & worry about \\
* & びんぼう & 貧乏 & poor \\
  & ペラペラ &      & fluent (in a language) \\
* & へん     & 変   & strange; unusual \\
* & らく     & 楽   & easy; comfortable \\
\multicolumn{4}{l}{\hlnotea{ウ --- 動詞}} \\
  & かぜがふく       & 風が吹く     & the wind blows \\
  & かつ             & 勝つ         & to win \\
  & コピーをとる     & コピーを取る & to make a photocopy \\
  & はこぶ           & 運ぶ         & to carry; to move \\
  & ひろう           & 拾う         & to pick up (something) \\
* & ほ（う）っておく & 放っておく   & to leave (someone/something/topic) \\
  &                  &              & along; to neglect \\
  & まにあう         & 間に合う     & to be on time \\
\multicolumn{4}{l}{\hlnotea{ル --- 動詞}} \\
  & けいかくをたてる & 計画を立てる & to make a plan \\
* & そだてる         & 育てる       & to raise; to bring up \\
  & たすける         & 助ける       & to help; to rescue \\
  & まける           & 負ける       & to lose (a match) \\
\multicolumn{4}{l}{\hlnotea{特別動詞}} \\
  & おねがいする & お願いする & to pray for help \\
  & さんせいする & 賛成する   & to agree \\
  & しっぱいする & 失敗する   & to fail; to be unsuccessful \\
  & てつやする   & 徹夜する   & to stay up all night \\
  & はんたいする & 反対する   & to oppose; to object to \\
  & ほんやくする & 翻訳する   & to translate \\
\multicolumn{4}{l}{\hlnotea{他の}} \\
* & こんなふう           &          & like this \\
* & じゆうに             & 自由に   & freely \\
* & ずっと               &          & for a long time; all the time \\
  & ぜったいに           & 絶対に   & definitely \\
* & そんなこと（は）ない &          & I don't think so. \\
  & 〜とか               &          & ... for example \\
* & なんども             & 何度も   & many times \\
* & もうすこし           & もう少し & a little more
\end{longtable}

% section tango_c22 (end)

\section{使役動詞 Causative Verbs}%
\label{sec:shiekidoushi_causative_verbs}
% section shiekidoushi_causative_verbs

The verb derivation that shall be introduced here is called the
\hldefn{causative form}. We use causative forms when we want to describe
who makes someone do something, or who lets someone do something.

\begin{table}[ht]
  \centering
  \caption{Rules of converting from base form to causative form}
  \label{table:rules_of_converting_from_base_form_to_causative_form}
  \begin{tabular}{l c l c l c l}
  \multicolumn{7}{l}{ル --- 動詞: 「る」は「させる」になる} \\
  食べる & $\to$ & 食べさせる &  &  &  &  \\
  $  $ \\
  \multicolumn{7}{l}{う --- 動詞: 最後の「う」の音は「あせる」になる} \\
  行く & $\to$ & 行かせる & & 話す & $\to$ & 話させる \\
  待つ & $\to$ & 待たせる & & 死ぬ & $\to$ & 死なせる \\
  読む & $\to$ & 読ませる & & 取る & $\to$ & 取らせる \\
  泳ぐ & $\to$ & 泳がせる & & 遊ぶ & $\to$ & 遊ばせる \\
  買う & $\to$ & 買わせる \\
  $  $ \\
  \multicolumn{7}{l}{特別動詞} \\
  来る & $\to$ & 来させる & & する & $\to$ & させる
  \end{tabular}
\end{table}

\begin{note}
  Any causative verb can be interpreted either with the
  ``\hlnotec{make} somebody do'' meaning, or the
  ``\hlnotec{let} somebody do'' meaning.
  We rely on our general knowledge about the situation
  to decide on whether the sentence connotes an authoritarian meaning
  or a supportive meaning.
\end{note}

\begin{eg}
  \begin{enumerate}
    \item お父さんは子供に野菜を食べさせました。
    \item 先生は私に英語を話させくれませんでした。
    \item 私は自分の子供に好きなことをさせてあげるつもりです。
    \item 私にこの仕事をやらせてください。
  \end{enumerate}
\end{eg}

% section shiekidoushi_causative_verbs (end)

\section{動詞＋なさい}%
\label{sec:doushi_nasai}
% section doushi_nasai

\begin{grammar}[動詞＋なさい]\index{動詞＋なさい}\label{grammar:doushi_nasai}
  When we mean to give a command in an authoritative manner,
  we use the form 動詞＋なさい.
\end{grammar}

\begin{note}
  なさい has a strong implication that one is ``talking down''
  to the person of which the command is given to.
  It is usually used by someone who, given the situation, believe
  that they are more mature, know better, and should be obeyed.
\end{note}

\begin{eg}
  \begin{enumerate}
    \item うちに毎日電話しなさい。
    \item かっこの中に単語を入れなさい。
    \item 文句を言うのをやめなさい。
  \end{enumerate}
\end{eg}

% section doushi_nasai (end)

\section{〜ば}%
\label{sec:ba}
% section ba

\begin{grammar}[〜ば]\index{〜ば}\label{grammar:_ba}
  We use \hlnoteb{〜ば} to express a condition at which an event that follows
  occur. We write
  \begin{center}
    A ば B
  \end{center}
  to say ``if A then B''.
\end{grammar}

\begin{table*}[ht]
  \centering
  \caption{Rules of conjugation for 〜ば}
  \label{table:rules_of_conjugation_for_ba}
  \begin{tabular}{l c l c l c l c l c l}
  \multicolumn{11}{l}{Verbs in the affirmative:} \\
  \multicolumn{11}{l}{「う」の音を捨てる、「えば」になる} \\
  食べる & $\to$ & 食べれば \\
  行く   & $\to$ & 行けば &  & 待つ & $\to$ & 待てば &  & 買う & $\to$ & 買えば \\
  する   & $\to$ & すれば \\
  来る   & $\to$ & 来れば \\
  $  $ \\
  \multicolumn{11}{l}{Verbs in the negative} \\
  \multicolumn{11}{l}{「い」を捨てる、「ければ」になる} \\
  行かない & $\to$ & 行かなければ
  \end{tabular}
\end{table*}

\begin{eg}
  \begin{enumerate}
    \item 車があれば、いろいろなところにいけます。
    \item 大家さんに言わなければ、わかりませんよ。
    \item 走れば、電車に間に合います。
    \item この薬を飲めば大丈夫です。
  \end{enumerate}
\end{eg}

% section ba (end)

\section{〜のに}%
\label{sec:noni}
% section noni

\begin{grammar}[〜のに]\index{〜のに}\label{grammar:_noni}
  We say ``A のに B'' to connote the meaning of ``despite A, B'',
  where both A and B are factual.
  The predicate in A is in the short form.
\end{grammar}

\begin{note}
  〜のに has a slight element of surprise to the sentence,
  in that one should expect B to not happen in the face of A,
  but B did happen.
\end{note}

\begin{eg}
  \begin{enumerate}
    \item この会社はお金があるのに、給料が安いです。
    \item 田中さんは親切なのに、山田さんは田中さんが嫌いです。
    \item 大きい問題なのに、あの人は誰にも相談しません。
  \end{enumerate}
\end{eg}

\begin{warning}
  We cannot use 〜のに for non-factual predicates; e.g.
  \begin{enumerate}
    \item $\times$ 日本語が少し難しいのに、これ読んでください。
    \item $\times$ あまり美味しそうじゃないのに、ここで食べましょう。
  \end{enumerate}
\end{warning}

% section noni (end)

\section{〜のような / 〜のように}%
\label{sec:noyouna_noyouni}
% section noyouna_noyouni

\begin{grammar}[〜のような / に]\index{〜のような / に}\label{grammar:_noyouna_ni}
  We say ``A のような B'' to mean B has the same quality or appearance
  as A.

  We say ``A のように B'' to mean B is done in the same way as A.
\end{grammar}

\begin{eg}
  \begin{enumerate}
    \item 私は鎌倉のような町が好きです。
    \item アウンサン・スーチーのような人になりたいです。
    \item メアリーさんは魚のように上手に泳げます。
  \end{enumerate}
\end{eg}

% section noyouna_noyouni (end)

% chapter dai_22_ka (end)

\chapter{第２３課}%
\label{chp:dai_23_ka}
% chapter dai_23_ka

\section{単語（２３課）}%
\label{sec:tango_c23}
% section tango_c23

\begin{longtable}{r l l l}
\multicolumn{4}{l}{\hlnotea{名詞}} \\
* & おもいで         & 思い出   & memory \\
  & かいがいりょこう & 海外旅行 & trip to a foreign country \\
  & からだ           & 体       & body \\
  & くつした         & 靴下     & socks \\
  & こくさいでんわ   & 国際電話 & international call \\
  & しゃかい         & 社会     & society \\
  & じゅぎょうりょう & 授業料   & tuition \\
  & しょうがっこう   & 小学校   & elementary school \\
  & せんきょ         & 選挙     & election \\
  & ソフト           &          & software \\
  & タイヤ           &          & tire \\
  & ただ             &          & free of charge \\
* & ばしょ           & 場所     & place \\
  & べっそう         & 別荘     & villa; vacation home \\
  & ボーナス         &          & bonus \\
* & ぼんおどり       & 盆踊り   & Bon dance (Japanese traditional dance) \\
  & めんせつ         & 面接     & interview \\
  & ゆうしょく       & 夕食     & dinner \\
  & りそう           & 理想     & ideal \\
  & るすばん         & 留守番   & looking after the house \\
  &                  &          & in the absence of someone \\
\multicolumn{4}{l}{\hlnotea{イ --- 形容詞}} \\
  & まずい           &          & (food / situation is) terrible \\
\multicolumn{4}{l}{\hlnotea{ウ --- 動詞}} \\
  & あめがやむ     & 雨が止む     & the rain stops \\
* & いなくなる     &              & (someone) is gone; to disappear \\
* & おせわにする   & お世話にする & to be in someone's care \\
* & おなかをこわす &              & to have a stomachache \\
  & きにいる       & 気に入る     & to find something agreeable \\
* & ちがう         & 違う         & to be different; wrong \\
  & なくなる       &              & to be lost; to disappear \\
  & わるぐちをいう & 悪口を言う   & to talk behind someone's back \\
\multicolumn{4}{l}{\hlnotea{ル --- 動詞}} \\
  & うける   & 受ける & to take (an exam, interview, etc.) \\
  & かえる   & 換える & to change \\
  & こたえる & 答える & to answer \\
  & はなれる & 離れる & (someone/something) separates; \\
  &          &        & parts from \\
\multicolumn{4}{l}{\hlnotea{特別動詞}} \\
* & 〜かおをする   & 〜顔をする & to look... (facial expression) \\
  & がっかりする   &            & to be disappointed \\
  & がまんする     & 我慢する   & to be tolerant/patient; to hold back \\
  & せわをする     & 世話をする & to take care of... \\
  & どうじょうする & 同情する   & to sympathize \\
  & パンクする     &            & (tire) goes flat \\
* & もどってくる   & 戻ってくる & (someone/something) comes back \\
  & ゆうしょうする & 優勝する   & to win a competition \\
\multicolumn{4}{l}{\hlnotea{他の}} \\
* & いや       &          & no \\
* & げんきでね & 元気でね & Take care of yourself. \\
  & さいごに   & 最後に   & finally \\
* & そうそう   &          & You are right. \\
* & そろそろ   &          & it is about time to... \\
* & そんな〜   &          & such...; that kind of... \\
* & ものそごく &          & extremely
\end{longtable}

% section tango_c23 (end)

\section{使役動詞の受動態 Causative-Passive Forms}%
\label{sec:shiekidoushi_no_jyudoudai_causative_passive_forms}
% section shiekidoushi_no_jyudoudai_causative_passive_forms

The ``causative-passive'' form is the passive form of causative sentences.

\begin{table}[ht]
  \centering
  \caption{Rules of converting from base form to causative-passive form}
  \label{table:rules_of_converting_from_base_form_to_causative_passive_form}
  \begin{tabular}{l c l c l c l}
  \multicolumn{7}{l}{ル --- 動詞: 「る」は「させられる」になる} \\
  食べる & $\to$ & 食べさせられる &  &  &  &  \\
  $  $ \\
  \multicolumn{7}{l}{う --- 動詞: 最後は「す」の「う」の音は「あせられる」になる} \\
  話す & $\to$ & 話させられる \\
  $  $ \\
  \multicolumn{7}{l}{う --- 動詞: 最後の「う」の音は「あされる」になる} \\
  書く & $\to$ & 書かされる & & 立つ & $\to$ & 立たされる \\
  読む & $\to$ & 読まされる & & 撮る & $\to$ & 取らされる \\
  泳ぐ & $\to$ & 泳がされる & & 遊ぶ & $\to$ & 遊ばされる \\
  買う & $\to$ & 買わされる \\
  $  $ \\
  \multicolumn{7}{l}{特別動詞} \\
  来る & $\to$ & こさせられる & & する & $\to$ & させられる
  \end{tabular}
\end{table}

\marginnote{The form 書かせられる is grammatically correct,
but it is less common than its counterpart 書かされる.}

\begin{eg}
  \begin{enumerate}
    \item 歌を歌わされました。
    \item 肉を食べさせられました。
    \item 私は友達に宿題を手伝わされました。
    \item ゆみはお母さんに勉強させられました。
  \end{enumerate}
\end{eg}


% section shiekidoushi_no_jyudoudai_causative_passive_forms (end)

\section{〜ても}%
\label{sec:temo}
% section temo

\begin{grammar}[〜ても]\index{〜ても}\label{grammar:_temo}
  We say ``A ても B'' to mean \hlnoteb{B even if A is true}.
\end{grammar}

\begin{note}[Comparison to 〜のに]
  In comparison to \hyperref[grammar:_noni]{〜のに},
  〜ても does not have the surprise element at all,
  and is usually said by a person who has a good understanding
  of why B happens despite A.

  Basically, 〜のに has the ``B happened despite A happening'' meaning,
  while 〜ても has the ``B is going to happen even if A happens'' meaning.
\end{note}

\begin{note}[Comparison to 〜たら]
  In comparison to 〜たら,
  which has a more straightforward if-then meaning,
  〜ても connotes the opposite of what the listener would expect
  in the event of A.
\end{note}

\begin{eg}
  \begin{enumerate}
    \item 雨が降っても、ピクニックに行きます。
    \item 暑くても、エアコンをつけません。
    \item 子供でも、わかります。
  \end{enumerate}
\end{eg}

\begin{ex}
  Try out the above examples by replacing 〜ても with
  〜のに and 〜たら, and justify the differences to yourself.
\end{ex}

% section temo (end)

\section{〜ことにする}%
\label{sec:kotonisuru}
% section kotonisuru

\begin{grammar}[〜ことにする]\index{〜ことにする}\label{grammar:_kotonisuru}
  We say ``A ことにする'', where A is a short form present tense verb,
  to mean that we have ``decided to do A''.
\end{grammar}

\begin{eg}
  \begin{enumerate}
    \item 車を買うことにしました。
    \item 今年の夏はベトナムに行くとこにしよう。
    \item 毎日十一時に寝ることにしています。
  \end{enumerate}
\end{eg}

% section kotonisuru (end)

\section{〜まで}%
\label{sec:made}
% section made

\begin{grammar}[〜まで]\index{〜まで}\label{grammar:_made}
  We say ``A まで'' to mean ``until A'', and we usually use this form
  to say that A is when some ongoing event will end or begin.
\end{grammar}

\begin{eg}
  \begin{enumerate}
    \item 晴れるまで、喫茶店でコーヒーを飲みながら、待ちます。
    \item 日本語は上手になるまで、国に帰りません。
    \item 赤ちゃんが寝るまで、（私は）歌を歌ってあげます。
    \item 日本の生活に慣れるまで、大変でした。
  \end{enumerate}
\end{eg}

% section made (end)

\section{〜方}%
\label{sec:kata}
% section kata

\begin{grammar}[〜方]\index{〜方}\label{grammar:_kata}
  Appending the noun-forming suffix 方 to the stem of a verb
  means the ``way of which the action is performed'' or ``how to do''.
\end{grammar}

\begin{eg}
  \begin{enumerate}
    \item 泳ぎ方 --- the way to swim; how to swim
    \item 考え方 --- the way to think; how to think
  \end{enumerate}
\end{eg}

\begin{eg}
  \begin{enumerate}
    \item 漢字を読む $\to$ 漢字の読み方
    \item 空港に行く $\to$ 空港の行き方 / 空港への行き方
    \item 日本語を勉強する $\to$ 日本語の勉強の仕方
  \end{enumerate}
\end{eg}

% section kata (end)

% chapter dai_23_ka (end)

\appendix

\chapter{Interesting Summaries as Exercise}%
\label{chp:interesting_summaries_as_exercise}
% chapter interesting_summaries_as_exercise

\begin{ex}
  Summarize to yourself the differences between the following if-then-like
  structures:
  \begin{center}
  〜ば \quad 〜ても \quad 〜なら \quad 〜たら \quad 〜のに \quad 〜と
  \end{center}
\end{ex}

\begin{ex}
  Collect the conjugation forms of passive, causative, and causative-passive,
  and find out the rules around causative-passive to conclude that
  the causative-passive form is simply as defined.
  \sidenote{This will help you in reducing a concept that you need to explicitly
  remember.}
\end{ex}

% chapter interesting_summaries_as_exercise (end)

\backmatter

\fancyhead[LE]{\thepage \enspace \textsl{\leftmark}}

% \nobibliography*
\bibliography{references}

\printindex

\end{document}
% vim:tw=80:fdm=syntax

