% !TEX TS-program = pdflatex
\documentclass[notoc,notitlepage]{tufte-book}
% \nonstopmode % uncomment to enable nonstopmode

\usepackage{classnotetitle}

\title{Archive of Problems for High School Mathematics}
\author{Johnson Ng}
\subtitle{A collection of problems for mathematics at the high school level}
\credentials{}
\institution{}

\input{latex-classnotes-preamble.tex}

\DeclareMathOperator{\RM}{RM}

\begin{document}
\makeatletter
\fancyhead[LE]{\thepage \enspace \textsl{\leftmark}}
\makeatother

\hypersetup{pageanchor=false}
\maketitle
\hypersetup{pageanchor=true}
\begin{fullwidth}
\chaptermark{Table of Contents}
\tableofcontents
\end{fullwidth}

\chapter*{Preface}%
\label{chp:preface}
\addcontentsline{toc}{chapter}{Preface}
% chapter preface

This archive comprises of a collection of problems in mathematics aimed at the
high school level. Questions may be from various sources, some of which I can
no longer identify for they were handed to me a long time ago as a print sheet.

I shall also note that the high school level of which I speak of is for the
Malaysian national education system. Readers from other nations should not
expect the level of this document to match the level of education in their own.

I am creating this archive with the purposes of collecting notable questions
that I believe are great questions for a learner, that they are able to open
their eyes to the topics that they are expected to learn in their time.

\section*{For Learners}

Marks are provided for each problem. Use them as a guide to suggest how much
work is expected to be put in, in the sense where if you've ended up producing
3 A4/letter papers of math to answer a problem worth only 4 marks, you're likely
thinking of the problem wrong, and you should go back to the drawing board.

The final answer for each problem, wherever sensible, will also be provided.
You've probably been told a million times in your life\sidenote{And if you
have not, you're hearing it here now.} that the final answer doesn't matter, but
what does is how you get there. Use the final answer to check for the accuracy
of your work.

If you do not get the correct final answer, chances are you've did something
wrong in your steps. Trace back, figure out what went wrong, whether it was a
silly mistake or a misconception, and rectify that. That's how you can learn
mathematics, and virtually any topic in any field with a clear structure.

% chapter preface (end)

\chapter{Inequalities}%
\label{chp:inequalities}
% chapter inequalities

(\textbf{1999, likely their SPM Trial})
\begin{enumerate}
  \item \begin{enumerate}
      \item Find the range of $x$ such that $9 + 2x > 3$ and $19 > 3x + 4$. [3m]
      \item Given that $2x + 3y = 6$, find the range of $x$ when $y < 5$. [2m]
    \end{enumerate}
    \marginnote{
      (\textbf{Answer})
      \begin{enumerate*}
        \item $-3 < x < 5$ \quad
        \item $x > \frac{-9}{2}$
      \end{enumerate*}
    }

  \item Find the range of values of $x$ if $(x - 2)(2x + 3) > (x - 2)(x +
    2)$. [3m]

    \marginnote{
      (\textbf{Answer}) $x < -1,\, x > 2$
    }
\end{enumerate}

% chapter inequalities (end)

\chapter{Logarithms}%
\label{chp:logarithms}
% chapter logarithms

(\textbf{1997, likely their SPM Trial})
\begin{enumerate}
  \item Show that $\log_{3} xy = 2 \log_{9} x + 2 \log_{9} y$. Hence find
    the value of $x$ and of $y$ which satisfy the equation $\log_{3} xy =
    10$ and $\frac{\log_9 x}{\log_9 y} = \frac{3}{2}$. [6m]

    \marginnote{
      (\textbf{Answer}) $y = 81, x = 729$
    }

  \item
    \begin{enumerate}
      \item Without using a calculator, find the values of $3^{\log_3 7}$. [2m]
      \item Solve the equation $5 \log_x 3 + 2 \log_x 2 - \log_x 324 = 4$ and
        give your answer correct to four significant figures. [4m]
    \end{enumerate}
    \marginnote{
      (\textbf{Answer})
      \begin{enumerate*}
        \item $7$
        \item $1.316$
      \end{enumerate*}
    }

  \item
    \begin{enumerate}
      \item Given that $2 \log_3 (x + y) = 2 + \log_3 x + \log_3 y$, show that
        $x^2 + y^2 = 7xy$. [3m]
      \item Without using a calculator, solve the equation
        \begin{equation*}
          \log_9 [ \log_3 (4x - 5) ] = \log_4 2.
        \end{equation*}
        [3m]
      \item After $n$ years, the value of a new car is $\RM 60 000 \left(
        \frac{7}{8} \right)^n$.  Calculate the number of years it takes for the
        value of the car to be less than $\RM 20000$ for the first time. [4m]
    \end{enumerate}
    \marginnote{
      (\textbf{Answer})
      \begin{enumerate*}
        \setcounter{enumii}{1}
        \item $8$
        \item $9$
      \end{enumerate*}
  }
\end{enumerate}

(\textbf{1998, likely their SPM Trial})
\begin{enumerate}
  \item Given that $\log_{\sqrt{x}} 4 = u$ and $\log_y 2 = w$, express $\log_4
    x^3 y$ in terms of $u$ and/or $w$. [4m]

    \marginnote{
      (\textbf{Answer}) $\frac{6}{u} + \frac{1}{2w}$
    }

  \item
    \begin{enumerate}
      \item Given that $\log_a 3 = x$ and $\log_a 5 = y$, express
        \begin{equation*}
          \log_a \left( \frac{45}{a} \right)
        \end{equation*}
        in terms of $x$ and $y$. [2m]
      \item Find the value of $\log_4 8 + \log_r \sqrt{r}$. [3m]
      \item Find the value of $x$ and of $y$ that satisfy the simultaneous
        equations $3(9^x) = 27^y$ and $\log_2 y = 2 + \log_2 (x - 2)$. [5m]
    \end{enumerate}
    \marginnote{
      (\textbf{Answer})
      \begin{enumerate}
        \item $2x + y - 3$
        \item $2$
        \item $x = \frac{5}{2},\, y = 2$.
      \end{enumerate}
    }
\end{enumerate}


(\textbf{1999, likely their SPM Trail})
\begin{enumerate}
  \item Given that $\log_2 3 = 1.585$ and $\log_2 5 = 2.322$, without using a
    calculator, find
    \begin{enumerate}
      \item $\log_2 45$ [3m]
      \item $\log_4 \left( \frac{9}{5} \right)$ [3m]
    \end{enumerate}
    \marginnote{
      (\textbf{Answer})
      \begin{enumerate*}
        \item $5.492$
        \item $0.424$
      \end{enumerate*}
    }

  \item 
    \begin{enumerate}
      \item If $x = \log_2 3$, find the value of $4^x$. Then, find the value of
        $4^y$ if $y = 1 + x$. [5m]
      \item Given that $\log_a 3 = 0.7924$,
        \begin{enumerate}
          \item prove that $\log_a 27a = 3.3772$, and
          \item solve the equation
            \begin{equation*}
              \sqrt{3} \times a^{n - 1} = 3
            \end{equation*}
            for $n$.
        \end{enumerate}
        [5m]
    \end{enumerate}
    \marginnote{
      (\textbf{Answer})
      \begin{enumerate*}
        \item $36$
        \item 
          \begin{enumerate*}
            \setcounter{enumiii}{1}
            \item $n = 1.3962$
          \end{enumerate*}
      \end{enumerate*}
    }
\end{enumerate}

(\textbf{2000, likely their SPM Trial})
\begin{enumerate}
  \item 
    \begin{enumerate}
      \item Solve $3^{\log_2 x} = 81$ for $x$. [3m]
      \item If $3^{2x} = 8(2^{3x})$, prove that $x \log_a \frac{9}{8} = \log_a
        8$. [3m]
    \end{enumerate}
    \marginnote{
      (\textbf{Answer})
      \begin{enumerate*}
        \item $x = 16$
      \end{enumerate*}
    }

  \item 
    \begin{enumerate}
      \item Simplify $\frac{\log_{12} 49 \times \log_{64} 12}{\log_{16} 7}$. [3m]
      \item If $3 \log_{10} xy^2 = 4 + 2 \log_{10} y - \log_{10} x$, and $x$ and
        $y$ are positive integers, show that $xy = 10$. [3m]
      \item The total amount of money deposited in a cooperative after a period
        of $n$ years is given by $\RM 2000 (1 + 0.07)^n$. Calculate the minimum
        number of year needed for the money to exceed $\RM 4000$. [3m]
    \end{enumerate}
    \marginnote{
      (\textbf{Answer})
      \begin{enumerate}
        \item $\frac{4}{3}$
        \setcounter{enumii}{2}
        \item $n = 11$
      \end{enumerate}
    }
\end{enumerate}

(\textbf{2001, likely their SPM Trail})
\begin{enumerate}
  \item Given that $\log_2 k = p$ and $\log_3 k =  r$, find $\log_k 18$ in terms
    of $p$ and $r$. [4m]
    \marginnote{
      (\textbf{Answer})
      $\frac{r + 2p}{pr}$
    }

  \item 
    \begin{enumerate}
      \item If $\log_{10} x = 2$ and $\log_{10} y = -1$, show that $xy - 100y^2
        = 9$. [3m]
      \item Solve
        \begin{enumerate}
          \item $3^{x + 2} = 24 + 3^x$, and
          \item $\log_3 x = log_9 (5x + 6)$.
        \end{enumerate}
        [3m]
    \end{enumerate}
    \marginnote{(\textbf{Answer})
      \begin{enumerate}
        \setcounter{enumii}{1}
        \item
          \begin{enumerate}
            \item $x = 1$
            \item $x = 6$
          \end{enumerate}
      \end{enumerate}
    }
\end{enumerate}

(\textbf{2002, likely their SPM Trail})
\begin{enumerate}
  \item 
    \begin{enumerate}
      \item Given $\log_5 3 = k$, and if $5^{2 \lambda - 1} = 15$, find
        $\lambda$ in terms of $k$.\sidenote{$\lambda$ is pronounced like
        lambda.} [3m]
      \item Solve
        \begin{equation*}
          \log_2 (7t - 2) - \log_2 2t = -1
        \end{equation*}
        for $t$. [3m]
    \end{enumerate}
    \marginnote{(\textbf{Answer})
    \begin{enumerate*}
      \item $\lambda = \frac{k+2}{2}$
      \item $t = \frac{1}{3}$
    \end{enumerate*}}

  \item 
    \begin{enumerate}
      \item Given $2 \log_4 x - 4 \log_{16} y = 3$, express $x$ in terms of $y$.
        [5m]
        \marginnote{(\textbf{Answer}) $x = 8y$}

      \item Solve the simultaneous equations
        \begin{equation*}
          2^{m - 1} \times 32^{k + 2} = 16 \quad\text{ and }\quad
          5^{-3m} \times 125^{3 - k} = 1
        \end{equation*}
        for $k$ and $m$. [5m]
        \marginnote{(\textbf{Answer}) $k = -2, m = 5$}
    \end{enumerate}
\end{enumerate}

% chapter logarithms (end)

\end{document}
% vim:tw=80:fdm=syntax

